\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-03-01 Draft_Form/tex/Discussion.tex
update 01_03_2018
update 01_03_2018

The aim of this study was to investigate the performance of design-based estimators for small area estimation of mean standing timber volume on two spatial forest management levels in Germany. It was of particular interest to gather information about the estimation error levels that can be attained using German NFI data that is characterized by low sampling intensities in the area of interests. To address these research questions, we applied the SRS, the \psmall{} and the \extpsynth{} estimators for cluster sampling to two forest management levels consisting of 45 and 405 small area units respectively in the German state of Rhineland-Palatinate.\par

>>>>

The aim of this study was to investigate the performance of model-assisted design-based estimators for small area estimation of mean standing timber volume on two spatial forest management levels in Germany. It was of particular interest to gather information about the estimation error levels that can be attained using German NFI data that is characterized by low sampling intensities in the area of interests. To address these research questions, we applied the SRS, the \psmall{} and the \extpsynth{} estimators for cluster sampling to two forest management levels consisting of 45 and 405 small area units respectively in the German state of Rhineland-Palatinate.\par

### 2018-03-01 Draft_Form/tex/Introduction(2).tex
update 01_03_2018
update 01_03_2018

Two-phase and three-phase samplings techniques have also been applied to small area estimation (SAE). SAE techniques address the situation where the number of samples within a subunit, or small area (SA), of the entire sampling frame is too small to provide reliable estimates for that unit. A broad range of SA estimators used in forest inventories \citep{kohl2006} originally comes from official statistics. One such method that is commonly applied is known as indirect estimation \citep{rao2015}, where statistical models are used to convert auxiliary information into predictions of the target variable that is rarely or not observed in the small area. These models are trained using data from outside the small area in order to "borrow strength" from areas where information is available. Of numerous applications of SAE in forestry \citep{breidenbach2012, goerndt2011, steinmann2013, mandallaz2013b}, most use unit-level models, i.e. the inventory plot is the unit of the response variable in the training data used for the model fit. Such unit-level models have been intensively investigated for timber volume estimation using various remote sensing auxiliary data \citep{koch2010, naesset2014inmaltamo}. Other studies have investigated area-level models, where the auxiliary information is only provided on the SA level \citep{magnussen2017}. Some studies have illustrated that even NFI data derived under low sampling densities can still be used to provide acceptable precision of small area estimates on much smaller management levels. One example is \citet{breidenbach2012} who used data from the Norwegian NFI to make small area estimation for standing timber volume for 14 municipalities where the number of NFI samples within these areas were between 1 and 35. The estimation errors under the applied model-based and design-based small area estimators turned to be markedly smaller than under the standard one-phase estimator. Another example is \citet{magnussen2014} who recently used the Swiss NFI data to estimate timber volume within 108 Swiss forest districts with sample sizes between 9 and 206. Similar studies using German NFI data for small area estimation have been lacking.\par

The aim of this study was to investigate whether the German NFI data can provide acceptable estimation precision on two forest district levels using the latest small area estimation procedures. We therefore conducted a study in the German federal state Rhineland-Palatinate where we extended the German NFI to a double-sampling design and applied three types of design-based small area regression estimators in order to derive point and variance estimates of mean standing timber volume for 45 and 405 forest districts respectively. The SA-estimators we considered were the \textit{pseudo-small}, \textit{extended pseudo-synthetic} and the \textit{pseudo-synthetic} design-based small area estimator suggested by \citet{mandallaz2013a} and \citet{ mandallaz2013b}. Auxiliary data consisted of a canopy height model (CHM) obtained from a countrywide airborne laser scanning (ALS) and a tree species classification map to be used for regression within tree species strata. The estimation precisions were compared to those obtained by the standard one-phase estimator for cluster sampling under simple random sampling. The chosen double-sampling estimators were selected for several reasons: \textbf{(i)} the design-based framework relaxes dependencies on the regression model assumptions which seemed appropriate facing severe quality restrictions in the ALS data; \textbf{(ii)} the estimators can be used with \textit{non-exhaustive}, i.e. non wall-to-wall, auxiliary information; \textbf{(iii)} all estimators are explicitly formulated for cluster sampling which has not yet been the case for frequently used model-dependent estimators; and \textbf{(iv)} the asymptotically unbiased g-weight variance accounts for estimating the regression coefficients on the same sample used for estimation (\textit{internal model approach}) and is also robust to heteroscedasticity of the model residuals. The results from this study were considered to provide valuable information whether the suggested procedure might be a cost-saving alternative to a regional FDI.\par

>>>>

Two-phase and three-phase samplings techniques have also been applied to small area estimation (SAE). SAE techniques address the situation where the number of samples within a subunit, or small area (SA), of the entire sampling frame is too small to provide reliable estimates for that unit. A broad range of SA estimators used in forest inventories \citep{kohl2006} originally comes from official statistics. One such method that is commonly applied is known as indirect estimation \citep{rao2015}, where statistical models are used to convert auxiliary information into predictions of the target variable that is rarely or not observed in the small area. These models are trained using data from outside the small area in order to "borrow strength" from areas where information is available. Of numerous applications of SAE in forestry \citep{breidenbach2012, goerndt2011, steinmann2013, mandallaz2013b}, most use unit-level models, i.e. the inventory plot is the unit of the response variable in the training data used for the model fit. Such unit-level models have been intensively investigated for timber volume estimation using various remote sensing auxiliary data \citep{koch2010, naesset2014inmaltamo}. Other studies have investigated area-level models, where the auxiliary information is only provided on the SA level \citep{magnussen2017}. Some studies have illustrated that even NFI data derived under low sampling densities can still be used to provide acceptable precision of small area estimates on much smaller management levels. One example is \citet{breidenbach2012} who used data from the Norwegian NFI to make small area estimation for standing timber volume for 14 municipalities where the number of NFI samples within these areas were between 1 and 35. The estimation errors under the applied model-dependent and design-based small area estimators turned to be markedly smaller than under the standard one-phase estimator. Another example is \citet{magnussen2014} who recently used the Swiss NFI data to estimate timber volume within 108 Swiss forest districts with sample sizes between 9 and 206. Similar studies using German NFI data for small area estimation have been lacking.\par

The aim of this study was to investigate whether the German NFI data can provide acceptable estimation precision on two forest district levels using the latest small area estimation procedures. We therefore conducted a study in the German federal state Rhineland-Palatinate where we extended the German NFI to a double-sampling design and applied three types of design-based small area regression estimators in order to derive point and variance estimates of mean standing timber volume for 45 and 405 forest districts respectively. The SA-estimators we considered were the \textit{pseudo-small}, \textit{extended pseudo-synthetic} and the \textit{pseudo-synthetic} design-based small area estimator suggested by \citet{mandallaz2013a} and \citet{ mandallaz2013b}. Auxiliary data consisted of a canopy height model (CHM) obtained from a countrywide airborne laser scanning (ALS) and a tree species classification map to be used for regression within tree species strata. The estimation precisions were compared to those obtained by the standard one-phase estimator for cluster sampling under simple random sampling. The chosen double-sampling estimators were selected for several reasons: \textbf{(i)} the design-based framework relaxes dependencies on the regression model assumptions which seemed appropriate facing severe quality restrictions in the ALS data; \textbf{(ii)} the estimators can be used with \textit{non-exhaustive}, i.e. non wall-to-wall, auxiliary information; \textbf{(iii)} all estimators are explicitly formulated for cluster sampling which has not yet been the case for frequently used model-dependent estimators; and \textbf{(iv)} the asymptotically unbiased g-weight variance accounts for estimating the regression coefficients on the same sample used for estimation (\textit{internal model approach}) and is also robust under heteroscedasticity of the model residuals. The results from this study were considered to provide valuable information whether the suggested procedure might be a cost-saving alternative to a regional FDI.\par

### 2018-03-01 Draft_Form/tex/Materials_and_Methods(2).tex
update 01_03_2018
update 01_03_2018

The estimators used in this study have been proposed by \citep{mandallaz2013a,mandallaz2013b} and derive their mathematical properties under the so-called infinite population approach. Therefore, we shall first provide a short introduction into this general estimation framework. We start by assuming that the population $P$ of trees $i \in 1,2, ..., N$ within a forest of interest $F$ is exactly defined, and each tree $i$ has a response variable $Y_i$ (e.g. its timber volume) that can be used to define the population mean $Y$ (e.g., the average timber volume per unit area) over $F$. Since a full census of all tree population individuals is almost never feasible, $Y$ has to be estimated based on a sample. In the infinite population approach this sample is a set of points or locations $x$ distributed independently and uniformly over the set of all possible points in $F$. Each point $x$ has an associated local density $Y(x)$ (e.g., the timber volume per unit area) whose spatial distribution is given by a fixed (i.e. non stochastic) piecewise constant function. The population mean $Y$ is mathematically equivalent to the integral of the local density function surface divided by the surface area of $F$, $\lambda(F)$, i.e. $Y=\frac{1}{N} \sum_{i=1}^{N}Y_i=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$, and thus the population mean $Y$ corresponds to a spatial mean. Since the actual local density function is unobserved in its entirety, one estimates $Y$ by taking a sample $s_2$ consisting of $n_2$ points and measuring each of their respective local densities. This sampling procedure is often referred to as \textit{one-phase sampling} (OPS) and $s_2$ is referred to as the terrestrial inventory. In contrast to the one-phase approach, \textit{two-phase} or \textit{double-sampling} procedures use information from two nested samples (phases). Practically speaking, the terrestrial inventory $s_2$ is embedded in a large phase $s_1$ comprising $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest $F$. For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some prediction model. The basic idea of this method is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed.

@@

@@

with $\bar{M}_{1,G}=\frac{\sum_{x \in s_{1,G}}M_{G}(x)}{n_{1,G}}$.\\

@@

\begin{equation}\label{ci_1phase}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
\hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_sae}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
\hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_global}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
\hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

In order to address the potential benefits of the small area estimators compared with the SRS approach, we calculated the \textit{relative efficiency} (Eq. \ref{eq:releff}) which can be interpreted as the relative sample size under SRS needed to achieve the variance under the double sampling estimators.

\begin{equation}\label{eq:releff}
rel.eff=\frac{\hat{\var}_{SRS}(\hat{Y})}{\hat{\var}_{SAE_{2phase}}(\hat{Y})}
\end{equation}

@@

% Graphik:
\begin{figure}[H]
	\centering
	\resizebox{0.9\hsize}{!}{\includegraphics*{fig/Auxvars_100dpi.png}}
	\caption{Left: CHM (top) and tree species classification map (bottom) available on the federal state level. Right: Magnified illustration of the supports used to derive the explanatory variables from the auxiliary data.}
	\label{fig:Auxvars}
\end{figure}

@@

The tree species map was used to predict the main tree species at each sample plot which served as an additional categorical variable called \textit{treespecies}. This involved two consecutive processing steps. In the first step, one of the five tree species was assigned to a sample location if 100\% of the raster values within the edge-corrected support were classified as that species. Otherwise, the sample location was assigned the value 'mixed'. Likewise for the CHM variables, the support radius was 12 meters although the use of different support sizes for each explanatory variable would be in agreement with the two-phase estimators presented in Section \ref{sec:SAestimators}. When using the \textit{treespecies} variable in a regression model, the support size and the percentage threshold parameters had to be optimized in order to minimize the variance within each level which subsequently leads to improved model precision. A detailed analysis and description of the optimal parameter processing for the explanatory variables of the present data set is provided in \citet{hill2017a}. In a second step, the \textit{treespecies} variable was also passed through a calibration model in order to reduce the effects of misclassification errors on the regression model coefficients and to increase model accuracy. The calibration model consisted of a decision tree from a random forest algorithm \citep{breiman2001} that was trained to predict the actual main plot tree species (known for all terrestrial plots) based on available auxiliary variables. These variables were the predicted \textit{treespecies} variable, the mean canopy height and standard deviation of the CHM, as well as the proportion of coniferous trees estimated from the classification map and the growing region derived from a polygon map. The algorithm was grown with 2000 trees considering 3 of the predictors for each split. We thus applied this calibration model to the \textit{treespecies} variable derived at all sample locations $s_1$. Table \ref{tab:classacc} gives the classification accuracies of the \textit{treespecies} variable after calibration.

>>>>

The estimators used in this study have been proposed by \citep{mandallaz2013a,mandallaz2013b} and derive their mathematical properties under the so-called infinite population approach. Therefore, we shall first provide a short introduction into this general estimation framework. We start by assuming that the population $P$ of trees $i \in 1,2, ..., N$ within a forest of interest $F$ is exactly defined, and each tree $i$ has a response variable $Y_i$ (e.g. its timber volume) that can be used to define the population mean $Y$ (e.g., the average timber volume per unit area) over $F$. Since a full census of all tree population individuals is almost never feasible, $Y$ has to be estimated based on a sample. In the infinite population approach this sample is a set of points or locations $x$ distributed independently and uniformly over the set of all possible points in $F$. Each point $x$ has an associated local density $Y(x)$ (e.g., the timber volume per unit area) whose spatial distribution is given by a fixed (i.e. non stochastic) piecewise constant function. The population mean $Y$ is mathematically equivalent to the integral of the local density function surface divided by the surface area of $F$, $\lambda(F)$, i.e. $Y=\frac{1}{N} \sum_{i=1}^{N}Y_i=\frac{1}{\lambda(F)}\int_{F}Y(x)dx$, and thus the population mean $Y$ corresponds to a spatial mean. Since the actual local density function is unobserved in its entirety, one estimates $Y$ by taking a sample $s_2$ consisting of $n_2$ points and measuring each of their respective local densities. This sampling procedure is often referred to as \textit{one-phase sampling} (OPS) and $s_2$ is referred to as the terrestrial inventory. In contrast to the one-phase approach, \textit{two-phase} or \textit{double-sampling} procedures use information from two nested samples (phases). Practically speaking, the terrestrial inventory $s_2$ is embedded in a large phase $s_1$ comprising $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^{\top}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest $F$. For every $x \in s_1$, $\pmb{Z}(x)$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using the choice of some prediction model. The basic idea of this method is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $s_1$ and to correct any possible model bias, i.e., $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the value of $Y(x)$ is observed. In this context, it is also important to note that the response and auxiliary variables are assumed to be error-free and the resulting errors for the point estimates reflect only the uncertainty due to sampling.

@@

\noindent with $\bar{M}_{2,G}=\frac{\sum_{x \in s_{2,G}}M_{G}(x)}{n_{2,G}}$.\\

@@

\noindent with $\bar{M}_{1,G}=\frac{\sum_{x \in s_{1,G}}M_{G}(x)}{n_{1,G}}$.\\

@@

%\begin{equation}\label{ci_1phase}
%CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
%\hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
%\end{equation}

\begin{equation}\label{ci_1phase}
CI_{1-\alpha}(\hat{\bar{Y}}_c)=\hat{\bar{Y}}_c \pm t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{\bar{Y}}_c)}
\end{equation}

@@

%\begin{equation}\label{ci_2phase_sae}
%CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
%\hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
%\end{equation}

\begin{subequations}\label{ci_2phase_sae}
	\begin{align}
CI_{1-\alpha}(\hat{Y}_{c,G,EXTPSYNTH})&=\hat{Y}_{c,G,EXTPSYNTH} \pm t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y}_{c,G,EXTPSYNTH})} \\
CI_{1-\alpha}(\hat{Y}_{c,G,PSMALL})&=\hat{Y}_{c,G,PSMALL} \pm t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y}_{c,G,PSMALL})}
	\end{align}
\end{subequations}

@@

%\begin{equation}\label{ci_2phase_global}
%CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
%\hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
%\end{equation}

\begin{equation}\label{ci_2phase_psynth}
CI_{1-\alpha}(\hat{Y}_{c,G,PSYNTH})=\hat{Y}_{c,G,PSYNTH} \pm t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y}_{c,G,PSYNTH})}
\end{equation}

@@

In order to address the potential benefits of the small area estimators compared with the SRS approach, we calculated the \textit{relative efficiency} (Eq. \ref{eq:releff}) which can be interpreted as the relative sample size under SRS needed to achieve the variance under the double-sampling (DS) estimators.

\begin{equation}\label{eq:releff}
rel.eff=\frac{\hat{\var}_{SRS}(\hat{Y})}{\hat{\var}_{DS}(\hat{Y})}
\end{equation}

@@

% Graphik:
\begin{figure}[H]
	\centering
	\resizebox{0.9\hsize}{!}{\includegraphics*{fig/Auxvars_400dpi.png}}
	\caption{Left: CHM (top) and tree species classification map (bottom) available on the federal state level. Right: Magnified illustration of the supports used to derive the explanatory variables from the auxiliary data.}
	\label{fig:Auxvars}
\end{figure}

@@

The tree species map was used to predict the main tree species at each sample plot which served as an additional categorical variable called \textit{treespecies}. This involved two consecutive processing steps. In the first step, one of the five tree species was assigned to a sample location if 100\% of the raster values within the edge-corrected support were classified as that species. Otherwise, the sample location was assigned the value 'mixed'. Likewise for the CHM variables, the support radius was 12 meters although the use of different support sizes for each explanatory variable would be in agreement with the two-phase estimators presented in Section \ref{sec:SAestimators}. When using the \textit{treespecies} variable in a regression model, the support size and the percentage threshold parameters had to be optimized in order to minimize the variance within each level which subsequently leads to improved model precision. A detailed analysis and description of the optimal parameter processing for the explanatory variables of the present data set is provided in \citet{hill2017a}. In a second step, the \textit{treespecies} variable was also passed through a calibration model in order to reduce the effects of misclassification errors on the regression model coefficients and to increase model accuracy. The calibration model consisted of a decision tree from a random forest algorithm \citep{breiman2001} that was trained to predict the actual main plot tree species (known for all terrestrial plots) based on available auxiliary variables. These variables were the predicted \textit{treespecies} variable, the mean canopy height and standard deviation of the CHM, as well as the proportion of coniferous trees estimated from the classification map and the growing region derived from a polygon map. The algorithm was grown with 2000 trees considering 3 of the predictors for each split. We thus applied this calibration model to the \textit{treespecies} variable derived at all sample locations $s_1$. Table \ref{tab:classacc} gives the classification accuracies \citep{congalton2008} of the \textit{treespecies} variable after calibration.
